Shedding Light on the ``Dark Side'' of Phylogenetic Comparative Methods

### 2015-12-08 Manuscript/response.tex

\documentclass[12pt]{letter}
\usepackage[a4paper,left=2.5cm, right=2.5cm, top=2.5cm, bottom=2.5cm]{geometry}
\usepackage{color}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage[osf]{mathpazo}
\usepackage{graphicx}
\usepackage{amsmath,amssymb}
\signature{Natalie Cooper \\ Gavin Thomas \\ Richard FitzJohn}
\address{Department of Life Sciences \\ Natural History Museum \\ Cromwell Road \\ London, SW7 5BD \\ natalie.cooper@nhm.ac.uk}
\longindentation=0pt
\begin{document}

@@

\textcolor{blue}{The reviewers and Associate Editor are positive about your work, however there are a few issues that need attention. Therefore, I invite you to respond to the reviewer(s)' comments and revise your manuscript. This might be closer to a Major Revisions in some ways: there are several issues to be addressed. At times the manuscript comes across as a bit whiny ("Why don't you lot do something?"). Tone can be difficult to get right, and there are some good suggestions (e.g. use Markdown to show your code), but after reading this I think I might avoid PCMs, as it seems there are a lot of pitfalls that I won't know about unless I work with someone who knows about them. Evidently it is not my destiny to come over to the Dark Side.}

We have tried to correct the tone where we can. It's very tricky!

\textcolor{blue}{I think the authors' suggestions that Someone Needs to Do Something would be stronger if they Started to Do Something. For example, they mention the BES QSIG's FGE pages, so why not start up a page on PCMs? An intro to running PCMs would be a necessary start, plus perhaps a Checklist of Things to be Checked (And How to Check Them). If the authors could find a day or two to set these up, then I think this would give the manuscript a more positive tone by announcing it, and might (if you are really really lucky) encourage others to add their knowledge.}

I'm involved with the FGE pages, and was planning to add a few PCM pages to this. As mentioned above, we had planned to "Do Something" and got a little way towards planning this with other members of the community, but personal constraints have prevented us from getting any further. We do hope to get something set up in the future, so have added a comment to this effect. We completely agree that this would be stronger with that kind of support set up already.

@@

Exactly how a song should be cited is not entirely clear but we have added a the following citation: Manic Street Preachers. 1996. Mr Carbohydrate. A Design for Life [CD 1], track 2. Columbia Records.

\textcolor{blue}{\textbf{Referee 1: Amy Zanne}}\\

\textcolor{blue}{As an audience member at this symposium and PCM end user, I was made well aware of the nature of use and abuse of PCMs. I walked away from this symposium oddly hopeful though that through collaborations, discussions and a number of items detailed in this review, solutions to intractable problems and public awareness of misuse can be addressed. I enjoyed this review as it provided concrete identification of the problems facing us without pointing fingers at anyone in particular. It also provided tractable solutions for how we can move forward. I found it a nice way to pause and take stock of where we are and how we can better go in the future.}

\textcolor{blue}{My biggest suggestion is that there can be some tightening of the text in a few places and a better mapping in others to the headings. I'll detail this specifically below.}

Thanks for the helpful and encouraging comments.

\textcolor{blue}{L179 I'd suggest that some of this folklore also comes from within the labs developing PCM. The reason Sam Price and I proposed the R hackathon at NESCent was because each of us had knowledge about PCM issues and solutions that were lab based from the groups where we came from who were developing methods. Some of this folklore is based on good practices, others not but the point is it is being passed down outside of the public literature.}

***We agree and make exactly these points on line XX: "Useful PCM folklore is often shared among developers, and among collaborating groups, but is not always shared outside of these circles." However to make this clearer we have specified that it "tends to be passed down from PIs to graduate students, and within labs developing these methods". We also have been more explicit that sometimes it is based on "best practice".

@@

\textcolor{blue}{L224-225 I think you can just say long and tedious or something along those lines. Why is “>20 pages” too long and what gives me a headache may be different to what gives you a “headache”.}

We have removed reference to headaches and also added the point that often equations are necessary.

\textcolor{blue}{L231 “This” is somewhat unclear. I'd make it clear that this review led to you needing to read more than 300 manuscript pages and a book to fully understand the issue.}

We have expanded on this to make the point that the difficulty relates to the problem of how much reading is necessary to have sufficient understanding.

\textcolor{blue}{L234 “required” for what? I'm assuming to correctly run the analysis or something along those lines?}

This section has been modified in response to Simon's comments so hopefully it is now clearer that we mean to understand the method.

@@

\textcolor{blue}{L257 drop a “to”}

Fixed.

\textcolor{blue}{L279 Again “incomprehensible” seems to depend on who is reading it. I'd argue that most methods papers are written for other methods developers as those are the audiences that will be reviewing the paper. They may or may not find the papers incomprehensible. The papers are likely to be less accessible but important for end users.}

****This is a very good point

\textcolor{blue}{L284 I'd go further than saying “there's no harm in doing”. In fact there's a lot of benefit (carrots). Although as you later note, we currently have no sticks making us do this. This seems like the best way to get the newly developed method into the hands of end users, make it accessible, but I agree that it is more work.}

Agreed, we have modified this line to say "there are great benefits to doing this"

\textcolor{blue}{L287 “help” with what? I fully agree that reproducibility is important in science but I'm unclear how it helps with this particular issue of “simplify, summarise and share”. I can imagine some avenues but I think it would be helpful to spell them out for the reader as otherwise it seems off topic.}

We have modified this to "Encouraging increased efforts at reproducibility would also help make methods more accessible to new users, by providing fully worked examples that can act as a starting point for their analyses.". Our point is that if someone uses a method and makes their analysis reproducible by sharing their code etc, new users can use this as a starting point for their own analyses.

\textcolor{blue}{L297 “bigger problem” for what? I'm guessing to come up to speed with understanding how the method works.}

Sorry we weren't clear here. The paragraph above is dealing with recommendations for new papers, but there are loads of already published papers that don't have accompanying blog posts or code etc., that also need to be summarised somehow. We have changed this sentence to "Summarising the glut of literature that already exists for established methods is a more difficult problem to solve."

\textcolor{blue}{L305 I've appreciated the Prometheus wiki ($http://prometheuswiki.publish.csiro.au/tiki-custom_home.php$) as a place for methods to be posted and reviewed, as well as providing an avenue for discussion of those methods. You have a number listed here so don't need another per se but I think this is a nice and well-used resource that would serve as a nice model for PCMs.}

It's always good to include more examples so we have added the Prometheus wiki to our list. We weren't aware of this beforehand so thanks for pointing it out.

\textcolor{blue}{L311 I think there's plenty of incentive but it's the carrot kind (it gets the PCM in an accessible form into the hands of an end user, meaning it's more likely to be used/reused) rather than the stick. Perhaps we do better with sticks than carrots, which is why Genbank works and more people are making public their data as journals require it?}

We have removed this line, but agree that a GenBank like system would be great for data sharing in ecology and evolution.

\textcolor{blue}{L313 To me this section reads as a series of suggestions rather than that we should be more cynical. I might broaden this subject heading. For instance, beginning with “It” on L323 this seems to be about best practices and things to keep in mind rather than being cynical. Basically you are outlining a nice checklist that we should keep in mind for most analyses (i.e., what's the sample size needed for enough power, do we need a complete and fully bifurcated phylogeny, etc.). Such a list would make a nice table… There's also of course assumptions/issues to consider that are very tied to specific PCMs. So, my take home is there are some things to keep in mind generally across most/all analyses and there are things to keep in mind specific to this analysis you are considering (the latter is harder to come by). There are a lot of “consider” and “for example” in this section.}

We have changed the subheading to "Factors to consider before using a method" to more accurately reflect its content. We have also rephrased the section to reduce the number of considers and for examples.

\textcolor{blue}{L341 It might also be useful to mention that reviewers and editors sometimes require authors to run PCMs even when these were not part of the original plan for the project. Sometimes these are useful suggestions that help to broaden the question, other times they aren't appropriate. It would be good to empower authors to say no when it's the latter.}

Agreed. This can be a big problem so we have may this point more explicitly: "In some cases, editors and reviewers may suggest using PCMs where they are not appropriate, and users should feel confident in rejecting these suggestions."

@@

\textcolor{blue}{L346 Again this section is more than just “Incentives” it feels perhaps like “Solutions” or something like that.}

We have changed this to "Solutions and incentives" to better reflect the content.

@@

\textcolor{blue}{L359 I agree that assumptions are not always given or are hard to find. Perhaps add “lists and” following “include”}

Added.

\textcolor{blue}{L364 I fully agree that it's hard to find funding for method development but it's also hard to find funding for simple data collection. I think what drives funding are theoretical/hypothesis driven science questions.}

Agreed. It's pretty hard these days to get funding for *any* kind of research! We have added a line "(although it is also hard to get funding for empirical projects)" to clarify that we're aware the situation is bad for everyone at present.

\textcolor{blue}{L366 I don't understand what this sentence means as written “This unfairly prejudices funding bodies (and hiring committees)”. Define “this” perhaps. I'd take it that funding bodies and hiring committees are creating these “biases” and driving them forward. We certainly need to change how the field appreciates methods developers but it comes at all of these levels.}

The point we were trying to make is that methods development is hard to get funded, thus your CV lacks big grants making it harder to secure faculty positions/tenure. however, we recognise that is was ambiguous, and also needs action at all levels. We have therefore deleted this line.

### 2015-12-08 Manuscript/dark.tex
Dealt with BOH and AZ's comments

Phylogenetic comparative methods are becoming increasingly popular for investigating evolutionary patterns and processes.
However, these methods are not infallible - they suffer from biases and make assumptions like all other statistical methods.
Unfortunately, although these limitations are generally well-known in the phylogenetic comparative methods community, they are often inadequately assessed in empirical studies leading to misinterpreted results and poor model fits.
Here we explore reasons for the communication gap dividing those developing new methods and those using them. 
We suggest that some important pieces of information are missing from the literature, and that others are difficult to extract from long, technical papers.
We also highlight problems with users jumping straight into software implementations of methods (e.g. in R) that may lack documentation on biases and assumptions that are mentioned in the original papers.
To help solve these problems we make a number of suggestions including providing
blog posts or videos to explain new methods in less technical terms, encouraging reproducibility and code sharing, making wiki-style pages summarising the literature on popular methods, more careful consideration and testing of whether a method is appropriate for a given question/dataset, increased collaboration, and a shift from publishing purely novel methods to publishing improvements to existing methods, and ways of detecting biases or testing model fit.
Many of these points are applicable across methods in ecology and evolution, not just phylogenetic comparative methods.

@@

  \subsubsection{1. Not everything is mentioned in the literature}
    As scientists we mainly communicate our ideas through the literature. 
    Unfortunately, some of the information needed to properly apply PCMs is not found in the literature. 
    We refer to this knowledge as ``PCM folklore'' because it tends to be passed down from PIs to graduate students (and it is occasionally closer to fiction than fact!).
    Sometimes the folklore includes tricks to get methods working, or useful rules-of-thumb; other times it is more opinion-based, but over time these opinions become rules. 
    Useful PCM folklore is often shared among developers, and among collaborating groups, but is not always shared outside of these circles. 
    When it is shared, it tends to be as email exchanges of ``dark advice'' that is not accessible to the rest of the community.

@@

  \subsection{2. The literature is too technical and/or important details are difficult to locate}
    Although some information is not found in the literature (see above), the majority of assumptions and biases of PCMs are documented somewhere. 
    A big issue for novice methods-users (and often for advanced users too) is that this information can be extremely technical and dense.
    It is not unusual for papers to be full of equations, \textgreater 20 pages long. 
    Such detail and mathematical literacy is critical in describing methods and facilitates testing/assessment, implementation and future developments of the method.
    In most cases equations could not simply be replaced with text.
   However, from the end user perspective it may more helpful to understand the assumptions and caveats.
   Within many comparative methods assumptions and caveats can be found in the Introduction, Methods, Results and/or Discussion of a paper. 
    They are rarely neatly corralled in one place.
     The sheer volume of literature can become a barrier to understanding, even, and perhaps especially, for the best known methods.
     For example, we recently reviewed papers discussing the assumptions and limitations of phylogenetic independent contrasts \citep{felsenstein1985phylogenies}, including how they are related to phylogenetic generalised least squares models \citep{garland2000using,rohle2006comment,blomberg2012independent}. 
    Even with prior knowledge of the key papers and authors to focus on, this resulted in $\approx 300$ manuscript pages and a book to read.
    The volume of reading itself is perhaps not the key issue, rather it is assessing when you have reached sufficient understanding which may be easy for experienced users but not clear for new users.    
    The combined effects of a vast literature and sometimes opaque assumptions make it easy to miss pertinent details.
    Many of these problems relate to due diligence for both reader and developer. 
    Methods developers are not responsible for making sure that all users read the literature. 
    The onus is, quite rightly in our opinion, on the reader/user to ensure that they have a clear understanding of the methods and caveats prior to using them. 
    However, simple steps, such as subheadings that point to caveats and assumptions, could add clarity and limit method misuse.

@@

    More recently however, the community has moved towards mostly implementing methods in R (\citealp{R-Core-Team:2014aa}; for a list of packages see Brian O'Meara's ``CRAN Task View: Phylogenetics, Especially Comparative Methods'' http://ftp.iitm.ac.in/cran/web/views/Phylogenetics.html), and code sharing has become almost ubiquitous. 
    The number of R packages for PCMs has increased markedly since 2005 when APE \citep{Paradis:2004aa} was released, and has increased particularly sharply since 2008 (Figure \ref{PCMRpackages}).
    Simultaneously, more people are able to use R thanks to workshops and changes in student training, thus when a new method is published it is now possible to take an R package ``off the shelf'' and use it to run the method immediately.
    The benefits of R are clear. 
    It is available to all, has a wide and engaged support community, and perhaps most importantly the source code for new methods is accessible.
    Users can, if they choose to, fully explore any new method by examining the source code and running their own simulations. 
    Indeed, \citet{freckleton2009seven} suggested that the ability to to conduct PCMs in flexible computing environments such as R would improve our ability to implement methods correctly. 
    However, this has instead led to an increasing number of people (including the authors of this paper) jumping straight to the implementation of a method, without fully understanding what the method is doing, what its assumptions might be, or what the results mean in a biological context.
    This is again an example of where users can and should conduct due diligence but also where developers and experienced users can help.
    Manuals, vignettes and help files for R packages rarely mention the assumptions of the method, or how to test model fit. 
    The APE book \citep{paradis2011analysis} for example, provides no guidance on assumption testing in its chapter on phylogenetic independent contrasts, even though the methods needed to do this are well-established and very easily implemented in R (for a counter example see CAIC and caper documentation; \citealp{purvis1995comparative,Orme:2013aa}).

\subsection{How can we solve some of these problems?}
  \subsubsection{1. Simplify, summarise and share}
    Many of the problems above are the result of methods that are difficult to communicate in purely descriptive terms.
    This is not entirely the fault of those writing these papers; most journals have strict word limits and a one line equation is generally a more succinct and precise descriptor of a method than a paragraph of text. 
    As we note above, the onus is on the reader to read extensively and to do due diligence.
    Nonetheless, due diligence is a responsibility at all levels of research and that includes ensuring clarity to all target readers, especially in describing assumptions and limitations.
    One solution is to prepare an accompanying blog post or video explaining the method in less technical terms. 
    Some journals already encourage this (including the British Ecological Society journals; see https://soundcloud.com/besjournals, https://www.youtube.com/user/MethodsEcolEvol), and even if they do not, there is no harm in doing this and hosting it on a personal website. 
    Some package developers provide blog updates (e.g. phytools http://blog.phytools.org) that serve this function. 
    The ability to share ideas with a non-technical audience is a key skill to develop, and may encourage more people to use the method.
    Encouraging increased efforts at reproducibility would also help, by including knitR \citep{Xie:2015aa} reports in Supplementary Material showing exactly how each analysis was run \citep[e.g.][]{fitzjohn2014much}, or by requiring all analyses and code to be available on GitHub or Bitbucket. 
    At the very least, having a list of the main assumptions, biases and caveats of the method somewhere obvious would reduce misuse and provide a place to point people to when they begin using a method (this will be difficult as there are often hidden assumptions in a method, and listing all possible assumptions and caveats may be unfeasible in some cases).
    An important point here is that we do not see these possible solutions as entirely the responsibility of the method developer or package writer. 
    The comparative methods community can help too and indeed often does via platforms such as r-sig-phylo.
    However, the ability to find helpful replies to lists depends on the subject headings and details in the text.
    A longer lasting solution would be to make published methods papers, and especially those that have a software focus, more readily updatable. 
    It should not be necessary to publish a new paper for updates of method.
    One example that gets around this problem is PLoS Currents where published papers can be updated relatively quickly with the full history of updated articles versioned.

    The glut of literature that already exists for established methods is a bigger problem.
    The Oxford Bibliographies Evolutionary biology pages have lists of key papers (http://www.oxfordbibliographies.com/obo/page/evolutionary-biology), but these are curated by just a few individuals and tend to contain a lot of papers. 
    One solution would be to establish a wiki-style website where people could post summaries of commonly used methods, along with lists of key papers to read. 
    The community would be responsible for peer reviewing these summaries to ensure all opinions are covered. 
    Erick Matsen's Phylobabble discourse page (http://phylobabble.org/) is an excellent example of a similar approach for phylogenetics. 
    The British Ecological Society's Quantitative Ecology Special Interest Group is also in the process of creating a ``Field Guide for Ecologists'' (http://bes-qsig.github.io/fge/) that will fill a similar niche in ecological methods.
    Unfortunately there is currently little incentive for people to contribute to such a project at the cost of time pursuing their own research.\\

  \subsubsection{2. Be more cynical about existing methods}
    A key skill to develop in science is cynicism, i.e., never take results from PCMs (or any other statistical analysis) at face value.
    As the Manic Street Preachers put it, ``Cynicism is the only thing that keeps me sane''.
    At a minimum, it is worth reading the original papers describing a method, plus any recent updates, and looking carefully for assumptions and caveats that may affect the analyses at hand.
    A good way to check a method is to simulate some data and see if the results are as expected \citep[e.g.][]{boettiger2012your}. 
    This can expose hidden assumptions or biases that have not been explored in the papers accompanying the method, or reveal a lack of understanding of the the mechanics of the method being used. 
    It is also important to consider whether the method will work on a particular dataset.
    For example it is worth considering how many species are required for reasonable power. 
    Often methods require more species than are usually  available. For example, the new trait-dependent diversification method of \citet{rabosky2015robust} is ``primarily applicable to phylogenies that include at least several thousand tips'' although the authors suggest that most empirical analyses have \textless 1000 tips. 
    Other considerations include whether the method is influenced by polytomies and whether the method is applicable to both ultrametric and non-ultrametric trees. 
    Indeed, many methods arbitrarily resolve polytomies using zero length branches, thus polytomies can inflate rates of evolution, and bias models of evolution  \citep{cooper2010body}. 
    Some current implementations of the OU model should not be used with non-ultrametric trees (e.g. MOTMOT; \citealp{Thomas:2011aa}) because they are based on transforming the tree directly, rather than transforming the variance co-variance matrix. The problem is that where at least one tip does not survive to the present the expected covariances relating each of those two tips with any other tip in the tree are not identical. Worked examples and explanations are provided in \citep{slater2014correction}. Although this is not a problem with applying the OU model to non-ultrametric trees per se, it is an example of different implementations of a common model that some users may not be aware of.

    It is also important to avoid retrofitting questions to the newest methods, instead we should think carefully about the question, whether the method is appropriate for the question, and whether PCMs are needed at all \citep{westoby1995misinterpreting,losos2011seeing}.
    Finally, end-users should never be afraid to question standard practice, sometimes it is just PCM folklore.

  \subsubsection{3. Incentives}
    It is important to recognize that our ability to do rigorous quantitative science often relies on highly-skilled methods-developers, especially as evolutionary biology becomes ever more computationally intensive. 
    We cannot afford to lose these people to industry, nor can we afford to pay industry wages, thus we need to make it worthwhile for such skilled researchers to remain in (or at least interact with) academia.

    First we need to stop insisting that methods papers are entirely novel. 
    Improvements to existing methods, and ways of detecting biases or testing model fit should be sufficient for publication.
    This is fairly standard in other fields, e.g. statistical phylogenetics, and these kinds of papers are arguably more useful to the community than constantly publishing new methods.
    When novel methods are published, journals should encourage researchers to include ways of testing the assumptions of their methods within the original publications and packages, and request simplified summaries to accompany technical papers.

    Second we need to fund pure methods development, including incremental methods. 
    Currently it is difficult to get funding for purely methods driven research; an empirical component is generally needed and methods development is often seen as part of the bigger empirical picture, rather than the reverse. 
    This unfairly prejudices funding bodies (and hiring committees) away from this kind of research.

    Finally, an obvious solution to many of these problems is for methods-developers and end-users to collaborate more. 
    Both parties can benefit greatly by collaboration. 
    Some benefits are obvious, e.g. methods-developers can gain extra datasets to test their ideas on and people who will discover corner cases and bugs in their software before it get released, whereas end-users can work with the most cutting-edge methods and software.  
    Most methods are not designed in a vacuum, they have a specific purpose usually based on predictions from theory, experimentation or observation.
    There is a huge benefit in sharing ideas as well as products (data and code) and time, as long as the benefits to both parties are not heavily asymmetric.

  \subsection{Conclusion}
    We are currently in an exciting period for phylogenetic comparative methods research.
    New methods are being published with increasing regularity, and we are also beginning to question older methods and classical ways of looking at comparative questions. 
    In addition, the field is becoming more open, with code being shared before analyses are even submitted for publication, and collaborative software development across groups, and even continents, becoming more common.
    However, while embracing these changes, we also need to ensure that we do not forget that PCMs have assumptions, caveats and biases like every other method. 
    These need to be highlighted so they can be accounted for in empirical analyses, and we need to be more active at providing ways of assessing these issues when publishing new methods.
    As members of the phylogenetic comparative methods community, we have a responsibility to find innovative ways to tackle these challenges.

>>>>

\begin{enumerate}
  \item Phylogenetic comparative methods are becoming increasingly popular for investigating evolutionary patterns and processes.
  However, these methods are not infallible - they suffer from biases and make assumptions like all other statistical methods.
  \item Unfortunately, although these limitations are generally well-known in the phylogenetic comparative methods community, they are often inadequately assessed in empirical studies leading to misinterpreted results and poor model fits.
  Here we explore reasons for the communication gap dividing those developing new methods and those using them. 
  \item We suggest that some important pieces of information are missing from the literature, and that others are difficult to extract from long, technical papers.
  We also highlight problems with users jumping straight into software implementations of methods (e.g. in R) that may lack documentation on biases and assumptions that are mentioned in the original papers.
  \item To help solve these problems we make a number of suggestions including providing
  blog posts or videos to explain new methods in less technical terms, encouraging reproducibility and code sharing, making wiki-style pages summarising the literature on popular methods, more careful consideration and testing of whether a method is appropriate for a given question/dataset, increased collaboration, and a shift from publishing purely novel methods to publishing improvements to existing methods, and ways of detecting biases or testing model fit.
  Many of these points are applicable across methods in ecology and evolution, not just phylogenetic comparative methods.
\end{enumerate}

@@

  \subsubsection{1. Not everything is mentioned in the literature}
    As scientists we mainly communicate our ideas through the literature. 
    Unfortunately, some of the information needed to properly apply PCMs is not found in the literature. 
    We refer to this knowledge as ``PCM folklore'' because it tends to be passed down from PIs to graduate students, and within labs developing these methods (and it is occasionally closer to fiction than fact!).
    Sometimes the folklore is based on best practice and includes tricks to get methods working, or useful rules-of-thumb; other times it is more opinion-based, but over time these opinions become rules. 
    Useful PCM folklore is often shared among developers, and among collaborating groups, but is not always shared outside of these circles. 
    When it is shared, it tends to be as email exchanges of ``dark advice'' that is not accessible to the rest of the community.

@@

  \subsection{2. The literature is too technical and/or important details are difficult to locate}
    Although some information is not found in the literature (see above), the majority of assumptions and biases of PCMs are documented somewhere. 
    A big issue for novice methods-users (and often for advanced users too) is that this information can be extremely technical and dense.
    It is not unusual for papers to be long, tedious and full of equations. 
    Of course, such detail and mathematical literacy is critical in describing methods and facilitates testing/assessment, implementation and future developments of the method.
    In most cases equations could not simply be replaced with text.
   However, from the end user perspective it may more helpful to understand the assumptions and caveats.
   Within many comparative methods papers, assumptions and caveats can be found in the Introduction, Methods, Results and/or Discussion of a paper. 
    They are rarely neatly corralled in one place.
     The sheer volume of literature can become a barrier to understanding, even, and perhaps especially, for the best known methods.
     For example, we recently reviewed papers discussing the assumptions and limitations of phylogenetic independent contrasts \citep{felsenstein1985phylogenies}, including how they are related to phylogenetic generalised least squares models \citep{garland2000using,rohle2006comment,blomberg2012independent}. 
    Even with prior knowledge of the key papers and authors to focus on, this resulted in $\approx 300$ manuscript pages and a book to read.
    The volume of reading itself is perhaps not the key issue, rather it is assessing when you have reached sufficient understanding which may be easy for experienced users but not clear for new users.    
    The combined effects of a vast literature and sometimes opaque assumptions make it easy to miss pertinent details.
    Many of these problems relate to due diligence for both reader and developer. 
    Methods developers are not responsible for making sure that all users read the literature. 
    The onus is, quite rightly in our opinion, on the reader/user to ensure that they have a clear understanding of the methods and caveats prior to using them. 
    However, simple steps, such as subheadings that point to caveats and assumptions, could add clarity and limit method misuse.

@@

    More recently however, the community has moved towards mostly implementing methods in R (\citealp{R-Core-Team:2014aa}; for a list of packages see Brian O'Meara's ``CRAN Task View: Phylogenetics, Especially Comparative Methods'' http://ftp.iitm.ac.in/cran/web/views/Phylogenetics.html), and code sharing has become almost ubiquitous. 
    The number of R packages for PCMs has increased markedly since 2005 when APE \citep{Paradis:2004aa} was released, and has increased particularly sharply since 2008 (Figure \ref{PCMRpackages}).
    Simultaneously, more people are able to use R thanks to workshops and changes in student training, thus when a new method is published it is now possible to take an R package ``off the shelf'' and use it to run the method immediately.
    The benefits of R are clear. 
    It is available to all, has a wide and engaged support community, and perhaps most importantly the source code for new methods is accessible.
    Users can, if they choose to, fully explore any new method by examining the source code and running their own simulations. 
    Indeed, \citet{freckleton2009seven} suggested that the ability to conduct PCMs in flexible computing environments such as R would improve our ability to implement methods correctly. 
    However, this has instead led to an increasing number of people (including the authors of this paper) jumping straight to the implementation of a method, without fully understanding what the method is doing, what its assumptions might be, or what the results mean in a biological context.
    This is again an example of where users can and should conduct due diligence but also where developers and experienced users can help.
    Manuals, vignettes and help files for R packages rarely mention the assumptions of the method, or how to test model fit. 
    The APE book \citep{paradis2011analysis} for example, provides no guidance on assumption testing in its chapter on phylogenetic independent contrasts, even though the methods needed to do this are well-established and very easily implemented in R (for a counter example see CAIC and caper documentation; \citealp{purvis1995comparative,Orme:2013aa}).

\subsection{How can we solve some of these problems?}
  \subsubsection{1. Simplify, summarise and share}
    Many of the problems above are the result of methods that are difficult to communicate in purely descriptive terms.
    This is not entirely the fault of those writing these papers; most journals have strict word limits and a one line equation is generally a more succinct and precise descriptor of a method than a paragraph of text. 
    As we note above, the onus is on the reader to read extensively and to do due diligence.
    Nonetheless, due diligence is a responsibility at all levels of research and that includes ensuring clarity to all target readers, especially in describing assumptions and limitations.
    One solution is to prepare an accompanying blog post or video explaining the method in less technical terms. 
    Some journals already encourage this (including the British Ecological Society journals; see https://soundcloud.com/besjournals, https://www.youtube.com/user/MethodsEcolEvol), and even if they do not, there are great benefits to doing this anyway and hosting it on a personal website. 
    Some package developers provide blog updates (e.g. phytools http://blog.phytools.org) that serve this function. 
    The ability to share ideas with a non-technical audience is a key skill to develop, and may encourage more people to use the method.
    Encouraging increased efforts at reproducibility would also help make methods more accessible to new users, by providing fully worked examples that can act as a starting point for their analyses.
    Reproducibility can be achieved by including knitR \citep{Xie:2015aa} reports in Supplementary Material showing exactly how each analysis was run \citep[e.g.][]{fitzjohn2014much}, or by requiring all analyses and code to be available on GitHub or Bitbucket. 
    At the very least, having a list of the main assumptions, biases and caveats of the method somewhere obvious would reduce misuse and provide a place to point people to when they begin using a method (this will be difficult as there are often hidden assumptions in a method, and listing all possible assumptions and caveats may be unfeasible in some cases).
    An important point here is that we do not see these possible solutions as entirely the responsibility of the method developer or package writer. 
    The comparative methods community can help too and indeed often does via platforms such as r-sig-phylo.
    However, the ability to find helpful replies to lists depends on the subject headings and details in the text.
    A longer lasting solution would be to make published methods papers, and especially those that have a software focus, more readily updatable. 
    It should not be necessary to publish a new paper for updates of method.
    One example that gets around this problem is PLoS Currents where published papers can be updated relatively quickly with the full history of updated articles versioned.

    Summarising the glut of literature that already exists for established methods is a more difficult problem to solve.
    The Oxford Bibliographies Evolutionary biology pages have lists of key papers (http://www.oxfordbibliographies.com/obo/page/evolutionary-biology), but these are curated by just a few individuals and tend to contain a lot of papers. 
    One solution would be to establish a wiki-style website where people could post summaries of commonly used methods, along with lists of key papers to read. 
    The community would be responsible for peer reviewing these summaries to ensure all opinions are covered. 
    Two excellent examples of a similar approach are Erick Matsen's Phylobabble discourse page (http://phylobabble.org/) for phylogenetics, and the Prometheus wiki for protocols in plant physiology (http://prometheuswiki.publish.csiro.au/tiki-custom\_home.php). 
    The British Ecological Society's Quantitative Ecology Special Interest Group is also in the process of creating a ``Field Guide for Ecologists'' (http://bes-qsig.github.io/fge/) that will fill a similar niche in ecological methods.
    We plan to establish a similar guide for PCMs in the near future with help from across the community.\\
    %Unfortunately there is currently little incentive for people to contribute to such a project at the cost of time pursuing their own research.\\

  \subsubsection{2. Factors to consider before using a method}
    A key skill to develop in science is cynicism, i.e., never take results from PCMs (or any other statistical analysis) at face value.
    As the Manic Street Preachers (\citeyear{manics}) put it, ``Cynicism is the only thing that keeps me sane''.
    At a minimum, users should read the original papers describing a method, plus any recent updates, and look carefully for assumptions and caveats that may affect the analyses at hand.
    A good way to check a method is to simulate some data and see if the results are as expected \citep[e.g.][]{boettiger2012your}. 
    This can expose hidden assumptions or biases that have not been explored in the papers accompanying the method, or reveal a lack of understanding of the the mechanics of the method being used. 
    It is also important to determine whether the method will work on a particular dataset.
    One key consideration is how many species are required for reasonable power. 
    Often methods require more species than are usually available. For example, the new trait-dependent diversification method of \citet{rabosky2015robust} is ``primarily applicable to phylogenies that include at least several thousand tips'' although the authors suggest that most empirical analyses have \textless 1000 tips. 
    Other considerations include whether the method is influenced by polytomies and whether the method is applicable to both ultrametric and non-ultrametric trees. 
    Indeed, many methods arbitrarily resolve polytomies using zero length branches, thus polytomies can inflate rates of evolution, and bias models of evolution  \citep{cooper2010body}. 
    Some current implementations of the OU model should not be used with non-ultrametric trees (e.g. MOTMOT; \citealp{Thomas:2011aa}) because they are based on transforming the tree directly, rather than transforming the variance co-variance matrix. 
    The problem is that where at least one tip does not survive to the present the expected covariances relating each of those two tips with any other tip in the tree are not identical. 
    Worked examples and explanations are provided in \citep{slater2014correction}. 
    Although this is not a problem with applying the OU model to non-ultrametric trees \textit{per se}, it is an example of different implementations of a common model that some users may not be aware of.

    It is also important to avoid retrofitting questions to the newest methods, instead we should think carefully about the question, whether the method is appropriate for the question, and whether PCMs are needed at all \citep{westoby1995misinterpreting,losos2011seeing}.
    In some cases, editors and reviewers may suggest using PCMs where they are not appropriate, and users should feel confident in rejecting these suggestions.
    Finally, end-users should never be afraid to question standard practice, sometimes it is just PCM folklore.

  \subsubsection{3. Solutions and incentives}
    It is important to recognize that our ability to do rigorous quantitative science often relies on highly-skilled methods-developers, especially as evolutionary biology becomes ever more computationally intensive. 
    We cannot afford to lose these people to industry, nor can we afford to pay industry wages, thus we need to make it worthwhile for such skilled researchers to remain in (or at least interact with) academia.

    First we need to stop insisting that methods papers are entirely novel. 
    Improvements to existing methods, and ways of detecting biases or testing model fit should be sufficient for publication.
    This is fairly standard in other fields, e.g. statistical phylogenetics, and these kinds of papers are arguably more useful to the community than constantly publishing new methods.
    When novel methods are published, journals should encourage researchers to include lists and ways of testing the assumptions of their methods within the original publications and packages, and request simplified summaries to accompany technical papers.

    Second we need to fund pure methods development, including incremental methods. 
    Currently it is difficult to get funding for purely methods driven research; an empirical component is generally needed and methods development is often seen as part of the bigger empirical picture, rather than the reverse (although it is also hard to get funding for empirical projects). 
    %This unfairly prejudices funding bodies (and hiring committees) away from this kind of research.

    Finally, an obvious solution to many of these problems is for methods-developers and end-users to collaborate more. 
    Both parties can benefit greatly by collaboration. 
    Some benefits are obvious, e.g. methods-developers can gain extra datasets to test their ideas on and people who will discover corner cases and bugs in their software before it gets released, whereas end-users can work with the most cutting-edge methods and software.  
    Most methods are not designed in a vacuum, they have a specific purpose usually based on predictions from theory, experimentation or observation.
    There is a huge benefit in sharing ideas as well as products (data and code) and time, as long as the benefits to both parties are not heavily asymmetric.

  \subsection{Conclusion}
    We are currently in an exciting period for phylogenetic comparative methods research.
    New methods are being published with increasing regularity, and we are also beginning to question older methods and classical ways of looking at comparative questions. 
    In addition, the field is becoming more open, with code being shared before analyses are even submitted for publication, and collaborative software development across groups, and even continents, is becoming more common.
    However, while embracing these changes, we also need to ensure that we do not forget that PCMs have assumptions, caveats and biases like every other method. 
    These need to be highlighted so they can be accounted for in empirical analyses, and we need to be more active at providing ways of assessing these issues when publishing new methods.
    As members of the phylogenetic comparative methods community, we have a responsibility to find innovative ways to tackle these challenges.
